Data Management for LSST Special Programs

### 2023-10-18 sec_proc.tex
dropped in details from sec:sci

% Rubin Observatory and the LSST Data Management System (DMS) are required 
% to process Special Programs data to produce unique and separate data 
% products whenever the original or reconfigured versions of the LSST 
% Science Pipelines can be used to do so (Section~\ref{sec:req}).

% The LSST Science Pipelines and the DMS are designed to process standard 
% (or alternative standard) visits.

% It depends on the science goals related to the Special Program, but some 
% Special Programs' data can and should be processed and served along with 
% (or as a part of) the WFD data products (Section~\ref{ssec:proc_wfd}).
% Others will require RDM to generate unique and separate data products 
% (Section~\ref{ssec:proc_rdm}).

% Cases where the Special Program obtains non-standard visits that cannot 
% be processed by the DMS (Section~\ref{ssec:proc_bounds}), and/or where 
% development of new algorithms or the allocation of significant additional 
% computational resources are needed, will require users to generate unique
% and separate data products (Section~\ref{ssec:proc_user}).

\subsection{Boundaries on Obtaining and Processing Non-Standard Visits} \label{ssec:proc_bounds}

Special Programs that do not use standard or alternative standard visits
might be limited by hardware or processing boundaries.

@@

A few potential challenges to Special Programs posed by hardware
boundaries are discussed below.

@@

The most likely challenges to Special Programs posed by processing
boundaries are discussed below.

@@

\subsection{Inclusion in the WFD Data Products}\label{ssec:proc_wfd}

Generally, data from Special Programs should be included in the Prompt 
and/or Data Release processing and data products alongside data from the 
WFD (i.e., included in the data products described by \citeds{lse-163}) 
whenever this is (1) possible and (2) scientifically beneficial.

Decisions about when and whether to include Special Programs data in 
the WFD's processing and data products are left to the discretion of the 
Rubin Operations Data Production and System Performance teams.

\subsubsection{WFD Prompt Processing and Alert Production}

@@

\subsubsection{WFD Solar System Processing}

@@

\subsubsection{WFD Data Release Processing}

@@

\subsection{RDM-Generated ``Unique and Separate" Data Products}\label{ssec:proc_rdm}

Rubin Data Management (RDM) is required to produce unique, separate, and 
joinable data products for Special Programs whenever this is possible with the 
original or reconfigured versions of the LSST Science Pipelines, and no new 
algorithmic development or significant computational resources are needed.

As described in Section~\ref{ssec:proc_wfd}, in some cases the science goals 
of a Special Program might be met by including the data in the WFD data 
products; especially the Prompt or DIA data products for standard visits.
Thus, most of the RDM-generated data products considered below are for 
static-sky Data Release data products, or data products for non-standard visits.
Decisions about when to produce separate Special Programs data products are 
left to the discretion of RDM.

@@

\subsection{User-Generated ``Unique and Separate" Data Products}\label{ssec:proc_user}

>>>>

A discussion of the anticipated details and challenges related to 
obtaining and processing data from Special Programs.

Appendix~\ref{sec:spcs} provides detailed examples for the processing
of data from Special Programs, including scenarios in which regular,
Special, and User-Generate Processing are all involved.

\subsection{Boundaries on Non-Standard Visits} \label{ssec:proc_bounds}

Special Programs that do not use standard or alternative standard visits
might be affected by hardware or processing boundaries.

@@

A few potential challenges posed by hardware
boundaries are summarized below.

@@

The most likely challenges posed by processing
boundaries are summarized below.

@@

\subsection{Regular (non-special) processing}\label{ssec:proc_reg}

Recall that the term ``regular processing" is used only within this document,
and refers to the image processing pipelines that are designed for, 
and will be applied to, the WFD program's observations (Section~\ref{sec:intro}).

Decisions about when to apply regular processing to Special Programs data,
or when to include it in the data products for the WFD program 
(e.g., if it improves the all-sky coadds), 
are ultimately left to the discretion of the Rubin Observatory's 
Data Management and System Performance teams.

\subsubsection{Prompt Processing and Alert Production}\label{sssec:proc_reg_prompt}

This might also include some non-standard visits (shorter or longer exposures), 
as long as they can be processed by the Prompt pipeline and an appropriate template image exists.
Visits with very short or very long exposure times (or very bright sky 
backgrounds) might be excluded if they would need specialized algorithms for,
e.g., instrument signature removal, difference-imaging, template-generation 
(Section~\ref{ssec:proc_bounds}).

If a Special Program's primary science goal requires specialized templates and 
Prompt processing, the Data Management System will have the capability to load 
and use an alternative template for some sky regions, based on the image metadata 
(i.e., the labels described in Section~\ref{ssec:sci_labels}). 
However, there would not be enough memory to hold alternative templates 
for the whole sky.

Including visits from Special Programs in regular Prompt processing alongside
vists from the WFD program is not, in general, anticipated to affect WFD science goals.
For example, analyses for a WFD-only subset could still be done using the program and
region labels described in Section~\ref{ssec:sci_labels}.

OLD

@@

\subsubsection{Solar System Processing}\label{sssec:proc_reg_ss}

@@

\subsubsection{Data Release Processing}\label{sssec:proc_reg_dr}

@@

\subsection{Special Processing}\label{ssec:proc_special}

As described in Section~\ref{ssec:sci_sproc}, 
Special Processing should be done by Rubin Data Management to 
produce unique and separate data products
for Special Programs when it is both possible and necessary.

In short, \emph{possible} means that original or reconfigured versions of the LSST
Science Pipelines can be used , and \emph{necessary} means the primary science goal for Special Program
could not be met without the data products produced by the Special Processing.
The definitions of possible and necessary, and the scope of Special Processing,
are ultimately left to the discretion of Rubin Data Management in Operations.

Specific examples are provided only to illustrate this interpretation of possible and necessary.
These examples do not place limits on what Special Processing might be done.

\item \textbf{Possible and necessary:}
in order to detect high-redshift (faint) galaxies in the DDFs,
Rubin Data Management uses the LSST Science Pipelines to deeply
coadd images for each field, and store the results of source
detection and characterization in unique and separate tables that
are included in the annual data release (Section~\ref{ssec:proc_rdm_ddf}).

\item \textbf{Possible but not necessary:}
a time-domain mini-survey that uses standard visits \emph{could}
have separate difference-image analysis object and source catalogs
generated, but this is not necessary as the science goals for the
mini-survey can be met by processing its data with regular Prompt
Processing (Section~\ref{ssec:sci_pproc}), and ensuring the
data is properly labeled (Section~\ref{ssec:sci_labels}).

\item \textbf{Possible but not necessary (secondary science goals):}
a time-domain mini-survey (or DDF) has a secondary science goal of detecting
precursor outbursts for transients, which requires coadding images
within windows of days, weeks, and months to reach various depths.
This set of custom coadds may be considered as ``not necessary" and requiring 
User-Generated Processing.

\item \textbf{Necessary but not possible:}
in order to find the most distant, faint Kuiper Belt Objects in the DDF,
a specialized, computationally intensive form of ``shift-and-stack" processing
is required for detection, but such algorithms are not used by the 
LSST Science Pipelines and so User-Generated Processing will be needed (Section~\ref{ssec:SPCS_TNO}).
A second example here is a twilight survey that uses non-standard visits 
which are outside the boundaries of what the LSST
Science Pipelines can process (Section~\ref{ssec:proc_bounds}).

\textbf{Tables should be joinable -- } where feasible.

@@

\subsection{User-Generated Processing}\label{ssec:proc_user}
